---
title: "Assignment 4 K-means for clustering"
author: "Avinash Ravipudi"
date: "2022-11-04"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("httr")
#install.packages("readr")
#install.packages("factoextra")
#install.packages("flexclust")
library(httr)
library(readr)
library(tidyverse)
library(factoextra)
library(ISLR)
library(flexclust)
library(caret)
```
#Importing Data set 
```{r}
#importing Data set and converting 
getwd()
pharma<-read.csv("Pharmaceuticals.csv")
#summarize the Data
#str(pharma)
head(pharma,10)
```
```{r}
set.seed(23)
#Data frame  Z Score scaling
pharma_scaled <- scale(pharma[,3:11])
summary(pharma_scaled)

# Data Frame Range Scaling 
pharma_range <- scale(pharma[,3:11])
```
The k-means algorithm was used to divide the 21 enterprises into three groups with no variable weights. We chose k=3 since that is the optimal k indicated by the silhouette approach.
```{r}
set.seed(23)
dst_rows <- get_dist(pharma_scaled)
fviz_dist(dst_rows) #To visualize distance between matrix rows
cluster1 <- kmeans(pharma_scaled, centers = 3, nstart = 15) # HEre taking K=3 & nstart=15
fviz_cluster(cluster1, data = pharma_scaled)
print(cluster1)
```


```{r}
fviz_nbclust(pharma_scaled, kmeans, method = "wss") # WSS method (ELBOW METHOOD)

fviz_nbclust(pharma_scaled, kmeans, method = "silhouette") #SILHOUETTE Methood (To find best K value)

```
I did not use the WSS approach since the graph did not show a distinct elbow and was extremely unclear. The graph does not indicate the elbow/knee position, and it flattens out more than once at k = 4 and 6, respectively, and I chose the silhouette approach since it is apparent to display the ideal cluster K = 5.

```{r}
#let's look at the mean value from actual data by clusters
aggregate(pharma[3:11], by=list(cluster=cluster1$cluster), mean) 
actual_data <- cbind(pharma, cluster = cluster1$cluster)
tibble(actual_data)
```
```{r}
by(actual_data, factor(actual_data$cluster), summary)#intensive statistical cluster analysis
```
Recommendations, Location and Exchange of cluster
```{r}
#Cluster median recommendation
T_Recom <- table(actual_data$cluster, actual_data$Median_Recommendation) 
names(dimnames(T_Recom)) <- c("Cluster", "Recommendation")
TR <- addmargins(T_Recom)
TR
```
The data do not show a clear link between clusterMedian Recommendation. There are 21 recommendations in total, with 1 strong buy, 7 moderate buys, 9 holds, and 4 moderate sells.
```{r}
#Cluster-based location breakdown
T_Location <- table(actual_data$cluster, actual_data$Location)
names(dimnames(T_Location)) <- c("Cluster", "Location")
Tlocation <- addmargins(T_Location)
Tlocation
```
We cannot deduce any association between cluster Location from the findings. A total of 21 firms are divided into 13 in the United States, three in the United Kingdom, and one each in Canada, France, Germany, Ireland, and Switzerland.
```{r}
#Exchange breakdown by cluster
T_Exchange <- table(actual_data$cluster, actual_data$Exchange)
names(dimnames(T_Exchange)) <- c("Cluster", "Exchange")
Texchange <- addmargins(T_Exchange)
Texchange
```
The results show that there is no link between clusterExchange. There are 21 corporations in all, divided into 1 Amex, 1 Nasdaq, and 19 NYSE.


Investigating the options
```{r}
fviz_nbclust(pharma_range, FUN = kmeans, method = "silhouette")
fviz_nbclust(pharma_range, kmeans, method = "wss")
```
We also perform tests to determine the best k using range normalization. The ideal k is 2 from the silhouette and 6 from the elbow (not clear). We'll stick with z-score normalization data because the k from range normalization isn't as good.`
```{r}
set.seed(11)
cluster2 = kcca(pharma_scaled, k=3, kccaFamily("kmeans"))
cluster2
clusters(cluster2) 

#Apply the predict() function
clusters_index <- predict(cluster2)
image(cluster2)
points(pharma_scaled, col=clusters_index, pch=17, cex=1.0)
```
To run kmeans cluster on k =3, we use the kcca algorithm instead of kmeans from basic R.

```{r}
set.seed(11)
cluster2 = kcca(pharma_scaled, k=3, kccaFamily("kmedians"))
cluster2
clusters(cluster2) 

#Apply the predict() function
clusters_index <- predict(cluster2)
image(cluster2)
points(pharma_scaled, col=clusters_index, pch=16, cex=1.0)
```


